### 2018-04-26 response.tex

\introcomment{~}{%
I agree with the peer reviewers that this paper addresses a topical subject of great importance to BSSA readers. The science is in the process of transitioning from purely observational based analysis of ground motions to a mixture of observations and simulations of ground motions. The paper is a useful contribution to this discussion, is generally well written, and carefully describes the methodology used and its limitations. The peer reviewers have made some very helpful suggestions for improving the paper that I hope the authors will strongly consider.
}

\response{%
Thank you. We were able to address all the suggestions made by the reviewers. In the vast majority of the cases we adopted their suggestions and we only disagreed with a very small portion of them, in which case we presented our arguments in the response to the reviews enclosed here.
}

\introcomment{~}{%
In particular, I agree with their comment that the limitations of the study, i.e., the choice of 11 metrics and earthquake simulations for only one earthquake, needs to be included in the abstract and the conclusions.}

\response{%
We modified the abstract and the conclusions accordingly. In the case of the abstract we added minor modifications to emphasize the use of 11 metrics, and a whole new sentence explicitly referencing the use of three simulation datasets corresponding to the 2008 Chino Hills earthquake. In the case of the conclusions, although we already had a paragraph describing the limitations of the study, we added a small sentence reemphasizing it.
}

\introcomment{~}{%
In addition, I also agree that all the terms and methodology should be well defined for the BSSA readership, so that the paper can be understood by them without resort to frequent online searches.
}

\response{%
\textcolor{red}{[ *** Pending ***]}
}

\introcomment{~}{%
I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit. It clearly shows that only these metrics are change significantly with the goodness of fit (P, F, G, E). Plotting these values as a function of P, F, G, and E together on the same figure might allow the reader to see this relationship more clearly and might help provide a heuristic understanding of how C4, C5, C6, and C8 are diagnostic whereas C11 is constant with P, F, G, and E, and thus not diagnostic. In fact, in my opinion, Figure 6 is a simple and easy way of determining which metrics are diagnostic that might even eliminate the need for the decision trees. In any case, more
discussion of this figure might help the authors make the case for the main conclusions.
}

\response{%
\textcolor{red}{[ *** Pending ***]}
% Thanks for the constructive suggestions. We agree with you and peer reviewers that the technical terms are not commonly used in seismology and strong motion earthquake engineering. The confusion happened because of integrating three different topics into a method to carry out this research. One technical term in one field is different in another field and that cause a confusion. For example cluster, class, and label all are referring to the same concept, however, it depends on the application. We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.
% \vspace{2ex}
% \newline
% We also modified Figure 6 in order to easily track the variation of different metrics for final classes. We do agree with you that Figure 6 clearly gives a good understanding about the metrics, however, we add decision trees, because they can convert the qualitative judgment to quantitative conditional form to be able to exactly decide on the final class based on their numerical GOF scores.
% \vspace{2ex}
% \newline
% We added text to abstract and conclusion to give a heads up to readers about 11 metrics and one earthquake with 3 velocity model  simulations.
}

@@

\response{%
We thank the reviewer for the positive feedback in this general review. We agree that there were some important points that we did not discuss fully in the manuscript and we have corrected them. In attention to this general review, we modified the abstract, the introduction, and the conclusions to (i) transparently point out the limitations of our study, and (ii) offer an additional perspective on the potential refinement of it or use with additional datasets or in reference to additional validation metrics.
}

\comment{1}{%
The terminologies used in the paper are not clear in many cases. For example, what are ``features'' ``attributes'' ``dimensions'' ``trees, nodes and leafs'' ``class''? It was very difficult to read the section on Data Analysis Method and Introduction since these terminologies were not clearly defined or described in terms of the variables used in this paper. I was confused what ``data'' and ``results'' were referring to: GOF score data? GOF metric data? GOF category? Something other than GOF? I suggest that the authors streamline the paper and the terminologies to be consistent and clear in the final version.
}

\response{%
\textcolor{red}{[ *** Pending ***]}
% We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering. The confusion happened because of integrating three different topics into a method to carry out this research. One technical term in one field is different in another field and that cause a confusion. For example cluster, class, and label all are referring to the same concept, however, it depends on the application. We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.
}

@@

\comment{2}{%
The terms ``common practice'' or ``common standards'' have been mentioned several times. Because GMSV is such a new field (and I don't think there is consensus, I for one don't agree with some statements but I agree that others may), I think these terms should be removed. Just say this is your opinion, or it is based on the references that you have listed.
}

\response{%
Thanks for the comment. We limited the use of these terms to a minimum and/or replaced them by sentences providing additional context.
}

\comment{3}{%
In abstract, ``semi-supervised'' and ``supervised'' learning techniques are mentioned, but in the paper these terms are not explained, what part of the learning techniques is semi-supervised and what does that mean?
}

\response{%
\textcolor{red}{[ *** Pending ***]}
% We briefly explained these terms \textbf{Data Analysis Method} section. We also added clarification about semi-supervised and supervised learning into \textbf{Introduction} section.
}

@@

\comment{4}{%
In introduction, the difference between ``verification'' and ``validation'' should be discussed.
}

\response{%
A brief description of the differences between verification and validation was added to the Introduction.
}

\comment{5}{%
In abstract and introduction, what does a ``direct application'' mean? it's not clear. Similarly, in Validation Metrics section: ``direct quantitative comparisons'' and ``indirect statistical analysis''.
}

\response{%
These terms were not present in the Abstract. The term ``direct application'' was present in the Introduction, but it has been removed due to improvements in the text based on other suggestion by this and a second Reviewer. The terms ``direct quantitative comparisons'' and ``indirect statistical analysis'' were present in the Introduction and the Validation Metrics sections. In our view, the first of these terms makes reference to validation through direct, i.e., signal to signal, quantitative comparisons of metrics; while the second of these terms makes reference to validation by means of statistical analysis showcasing the overall results of comparisons, and thus the use of ``indirect''. We understand, however, that this may not be a view shared by any other reader, so we have modified the text to be more explicit about this distinction.
}

@@

\comment{7}{%
The two questions posed in the introduction lines 59--61: I don't think the first one is really the question, all parameters should be considered if anyone has thought of them as important (i.e., has proposed them). The second question is the main question and the one being addressed in this paper.
}

\response{%
We thank the reviewer for expressing his/her opinion about this point. First, we changed slightly the sentences as a whole to emphasize these are indirect questions as opposed to direct questions. That said, we do think they both are valid and relevant questions. The argument made by the reviewer that any or all parameters should be included if someone has thought of them as important is precisely one of the points we argue against, and the results of this research support that argument. The fact that someone thinks of a parameter as relevant does not guarantee its relevance, it is only the reflect of his/her expert (though still subjective) opinion. There may be parameters that are more influential, and these can be as influential as to the point of proving others irrelevant. This is the advantage of the procedure we follow here, which allows one to evaluate objectively, based on how data relates to the results, how relevant or irrelevant is each one of the metrics involved in reaching the same conclusions. This, in turn, allows one to prioritize the metrics, which is the point of the second indirect question posed in this part of the introduction.
}

\comment{8}{%
You are treating these metrics as ``independent variables'', but as you have also noted many of them are not. For example, Arias intensity and energy are correlated. Does this compromise your conclusions? If not explain why in the Results section.
}
%
\response{%
\textcolor{red}{[ *** Pending ***]}
% This is a very interesting comment. You are totally right. There are fairly high correlation between Arias intensity and energy integral. However, decision tree algorithm is robust enough to realize this similarity and pick the best one in each decision making process. As we remove the peak ground acceleration and response spectra from the prediction model, the algorithm looks for other best parameters. In case of hundreds of metrics it would be a reasonable choice to do feature selection analysis and choose one metric out of several correlated metrics. However, in this study, we only deal with 11 metrics and the results show that it works fairly well.
}

@@

\response{%
We followed this suggestion and deleted the sentence.
}

\comment{10}{%
Page 14, line 310: you say ``...the latter can essentially be discarded...'' I don't understand this statement, wouldn't this mean that both C1 and C7 are important to consider since they are independent of each other (one gives information that the other one doesn't)?
}

\response{%
In the specific case of the 2-feature clustering in the C1-C7 subspace, what the results show is that only C7 (in the vertical axis) is playing a role in establishing boundaries between the different classes. C1 does not help to do that. C1 may be important in other context, but here it does not add any value to the decision making process. That is why we call it irrelevant feature, because one can define the clusters based on only defining boundaries on C7. It is normal to have a tendency to interpret this as a correlation problem, in which the interpretation of independence (i.e., lack of correlation between C1 and C7) would indicate that both are relevant. But that is not the goal of the clustering process. The goal is to identify which metrics play a role in delimiting the classes when correlated to other metrics. In other words, the colors and the way boundaries end up forming in these figures, is as important as the shape of the clouds of data-points.
}

\comment{11}{%
Page 14, line 323: It's not clear how you ``replicate data'', provide more detail.
}

\response{%
We modified this paragraph to provide more details about the oversampling process.
}

\comment{12}{%
I don't think ``confusion matrix'' given in tables is explained in the paper!
}

\response{%
We are a little surpirsed by this comment. The paper does have an explanation about confusion matrices. In the original manuscript, this explanation is the section titled \textbf{Decision Trees}, page 13, between lines 280 and 282. The explanation includes an illustrative example for the simplest case of two categories and it should not be difficult for any reader of the BSSA to understand this concept is extendable to larger matrics such as those shown in Tables 2 and 3. We added a sentence in the explanation to make it explicit that they can be extended to deal with more attributes.
}

@@

\response{%
You are right. It could be misleading. We reworded the second half of this paragraph to make things more transparent and to keep them in the context of what we used in our study.
}

\comment{14}{%
Page 18, line 405: this is implying that fig 10 is similar to fig2! I personally cannot see this in the figures! They look very different. It might be better to regenerate fig 11a but with different color scales that only vary by four choices, then the comparison would be much easier.
}

\response{%
In hindsight, we may have allowed ourselves to be carried away by our own optimism. We have therefore modified our comments in this section to a more self-critical approach about the results. The question is really not whether they look the same or not, but whether they lead to similar---if not equivalent---conclusions. It is our interpretation that they do in the sense that looking at the results obtained with either approach one may walk away with the same conclusion about the simulation as a whole and about how well (or not well) did the simulations turned out with respect to specific areas in the simulation domain. We modified the text to convey this message, now hopefully more clearly than before.
}

\comment{15}{%
Figure 8: I don't know what you mean by y-label ``decision attributes''. Do you mean ``the number of GOF metrics''? why there is no data point for 1 or 11?
}

\response{%
That is right. It should be the number of decision attributes. We modified the figure accordingly. The reason why there is no data for 1 is because there are no trees that would classify the data using a single metric into the four classes because these are binary trees, i.e., each node represents a binary decision, so one needs at least 2 metrics with 3 or more nodes with a minimum depth of 2. On the other end of things, the reason why there is no data for 11 metrics is because the analysis of the data resulted in the metric C11 not having any participation in the decision process, as shown in the original Table 6. In new version of Table 6 we include cases in which we forcefully removed metric C8, first, and then C5 and C8. Once one does that, the participations of the metrics changes and so would the topology of the trees.
}

\comment{16}{%
The conclusion basically says that response spectrum is the most important validation metric. This is not surprising to any engineer or modeler. The big question is what other parameters describe the differences between a simulation and a recorded motion if their response spectra matched? Or in other words, what are the other important validation parameters aside from response spectrum (many have argued that's duration, but no statistical proof). Can you use your method to answer this question? Can you condition your machine learning process on Sa being the same (for example use a subset of data with similar Sa or artificially make them equal), then decide what's the next important parameter? Would the result be the same as in your conclusions? Or would it be influenced by removal of Sa as a parameter (this is possible since there are correlations between Sa and other metrics).
}

\response{%
\textcolor{red}{[ *** Pending ***]}
% That is a very good suggestion. We excluded C8(response spectra) and C5(peak acceleration) from analysis and reproduce similar trees with same complexities in terms of pruning process with previous four trees. As we expected, most important, metrics are still C4 and C6.  C11 also become somewhat effective in the most complicated tree.
}

@@

\newpage

\begin{center}
	\bf
	\large
	Authors' Response to Reviewer 2
\end{center}

\noindent
We thank the reviewer for his/her comments and suggestions, which helped us improve the manuscript. To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font and the removed text in red font. In the following, we provide a response to each comment. The original comments are in italic black font, followed by our responses in regular blue font.
\vspace{2ex}
\newline

\introcomment{~}{%
Earthquake ground motion simulations are usually validated against observational data or empirically-derived solutions. However, exactly which parameters to compare (e.g., peak ground velocity, duration, spectral content) in order to assess the goodness-of-fit, is not always clear. This paper uses machine learning techniques and eleven possible metrics for comparing (recorded and simulated) ground motions to determine the parameters that are most predictive of goodness-of-fit. This is an interesting paper, and will help future researchers prioritize which metrics to focus on when validating ground motions simulations. The paper is generally well written.
}

\response{%
We thank the reviewer for the generally positive view of our paper.
}

\introcomment{~}{%
As a strong ground motion seismologist (i.e., I assume I'm part of the audience you hope to reach), there is a lot of machine learning-related jargon that makes certain sections difficult to follow.
}

\response{%
\textcolor{red}{[ *** Pending ***]}
% We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering. The confusion happened because of integrating three different topics into a method to carry out this research. One technical term in one field is different in another field and that cause a confusion. For example cluster, class, and label all are referring to the same concept, however, it depends on the application. We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.
}

\introcomment{~}{%
It would also be useful to have some discussion regarding which parameters were found to be repetitive, in case future authors wish to look at more than the four recommended parameters.
}

\response{%
\textcolor{red}{[ *** Pending ***]}
% We also studied the models without response spectra (C8) and without response spectra (C8) and peak ground acceleration (C5).
}

@@

\comment{2}{%
Line 11: ``...the objective \textbf{of}...''
}

\response{%
Added. Thanks.
}

\comment{3}{%
Lines 25--41: The frequent use of phrases such as ``Among these methods,''; ``Among the latter,''; ``In this category,''; and ``Within this group,'' make the text difficult to follow.
}

\response{%
We modified the second half of the first paragraph in the Introduction to be more direct in the description of the alternative verification and validation metrics.
}

@@

\comment{4}{%
Lines 42--43: I would suggest moving the description of the Anderson (2004) method (i.e., Line 48+) to directly follow this sentence.}

\response{%
Thanks for the suggestion. We moved it and it does read better this way.
}

\comment{5}{%
Lines 59--61, 391--394: It should be stated somewhere that this is really intended for \textbf{engineering purposes}.
}

\response{%
While it is true that the initial intent of Anderson (2004) was to develop credibility of synthetic seismograms for engineering applications, we do not share the need to make that intent explicit here. On the one hand, these metrics have already been used by several authors who did not intend to validate seismograms for any engineering application, but to simply show---by means of them---that their synthetics were realistic in a general sense. On the other hand, we do not want to convey the idea that the steps that we followed are exclusive to narrowing validation criteria for engineering applications. Others may see our work and these steps as applicable to other ends. That being said, we hope the reviewer will be pleased to notice that we modified the Introduction to point out that engineering applications are the one of the motivating factors for the increased attention in verification and validation. Please see the first sentence in the introduction (see line 29).}

@@

\comment{9}{%
Line 236: Where is the discussion on the consequences of overfitting? The only other place I see overfitting mentioned is line 328.
}

\response{%
We added it at this point. Thank you.
}

\comment{10}{%
Line 236: Remove ``A matter we discuss later.''
}

\response{%
We removed it, and in its place added an explanation about overfitting in response to the previous comment.
}

\comment{11}{%
Lines 241--243: It is not helpful to list different algorithm names for making decision trees, without any information about them or how they differ.
}

\response{%
We disagree about the need to make detailed distinctions about these different algorithms. That is why we provided references. If the reader is interested in learning the differences, the most we can do without expanding the scope of our study is to point him/her in the right direction, which we already do. Nonetheless, in attention to the Reviewer, we modified slightly the descritpion to give some additional context in what regards to the C5.0 algorithm and added a sentence explaining that detailing the alternative algorithms is out of the scope of this study.
}

@@

\comment{12}{%
Lines 257--258: I don't understand this sentence. What is a continuous attribute?
}

\response{%
Following Equation (3) we define $A[a]$ as all the possible values of attribute $A$. The use of $[a]$ implies that $A$ is considered a numerically discrete attribute. We therefore meant to say that if $A$ continues attribute. Thanks for the comment. Fixed.
}

@@

\comment{14}{%
Line 300: Isn't this also true for the C5-C7 combination?
}

\response{%
That is right. We modified the sentence to reflect that.
}

\comment{15}{%
Line 328: What is ``a strong pruning process''?
}

\response{%
We modified it to ``heavy pruning process'' and added an explanation earlier in the Decision Trees section.
}

### 2018-04-26 decision-tree.tex
close to finishing

Upon completing the clustering and resampling processes, the next step is to subdivide the dataset in two, a training dataset (with 70 percent of the samples picked randomly), and a testing dataset (with the remaining 30 percent). We first use the training dataset to build a large suite of potential decision trees, and then use the testing dataset to evaluate and pick the best possible decision tree(s) among them. There are several algorithms for building decision trees \citep[e.g., ID3, C4.5, C5.0, CART; see][]{Quinlan_1986_ML, Quinlan_1993_Book, Quinlan_1996_JAIR, Breiman_1984_Book}. \myrevision{C5.0 is considered as the developed version of ID3 and C4.5, where it does not have any limitations in handling missing data, numerical attributes, and developing boosted models. Internal algorithm of CART is different than other methods, however, discussing technical details of these algorithms are beyond the scope of this study.} Here we use the C5.0 implementation \citep{Kuhn_2017_Manual}, which is an extension of the C4.5 algorithm developed by \citet{Quinlan_1993_Book, Quinlan_1996_JAIR}.

@@

The effectiveness of the tree, on the other hand, needs to be measured. To that end we use the factor $F_{\beta}$ proposed by \citet{Rijsbergen_1979_Book} as
% 
\begin{equation}
	\label{eq:f}
	F_{\beta} = \frac{ (1 + \beta^2) P R}{\beta^2 P + R}
	\, ,
\end{equation}
% 
where $P$ and $R$ are the precision and recall factors, respectively, and $\beta$ is a weighting factor between the two. $P$ measures the fraction of data samples that are correctly classified within a category, whereas $R$ measures the fraction of the data classified within that category regardless of whether it was correctly classified or not. Mathematically, they are defined as
% 
\begin{equation}
	P = \frac{ \mathrm{TP} }{ \mathrm{TP} + \mathrm{FP} } 
	\quad \mathrm{and} \quad
	R = \frac{ \mathrm{TP} }{ \mathrm{TP} + \mathrm{FN} }
	\, .
\end{equation}
% 
Here, TP, FP, and FN are the number of samples in the testing dataset considered to be true-positives, false-positives, and false-negatives, respectively. These values are typically ordered in what is known as a confusion matrix. For the simplest case of two categories, a confusion matrix takes the form
% 
\begin{equation}
\nonumber
\begin{array}{rr@{}rl}
	&					&	\multicolumn{2}{c}{\mathrm{Prediction}}		\\
	&					&	\mathrm{Positive}	&	\mathrm{Negative}	\\[0.75ex]
	\mathrm{Actual~Value}		
	&	\begin{array}{r}
			\mathrm{Positive} \\[1ex]
			\mathrm{Negative} 
		\end{array}
	&	\left[~
		\begin{array}{c}
			\mathrm{TP} \\[1ex]
			\mathrm{FP} 
		\end{array}\right.
	&
		\left.
		\begin{array}{cc}
			~\mathrm{FN} \\[1ex]
			~\mathrm{TN} \\[1ex]
		\end{array}~\right]
\end{array}
\end{equation}
% 
where TN is the number of true-negative samples.

>>>>

Upon completing the clustering and resampling processes, the next step is to subdivide the dataset in two, a training dataset (with 70 percent of the samples picked randomly), and a testing dataset (with the remaining 30 percent). We first use the training dataset to build a large suite of potential decision trees, and then use the testing dataset to evaluate and pick the best possible decision tree(s). There are several algorithms for building decision trees \citep[e.g., ID3, C4.5, C5.0, CART; see][]{Quinlan_1986_ML, Quinlan_1993_Book, Quinlan_1996_JAIR, Breiman_1984_Book}. \change{Here, we use the C5.0 algorithm as implemented by \citet{Kuhn_2017_Manual}. This algorithm is the latest update to the original ID3 and subsequent C4.5 algorithms \citep{Quinlan_1993_Book, Quinlan_1996_JAIR}. C5.0 is superior to its predecessors because it reduces the limitations for handling numerical attributes, missing data, and it has additional features such as the development of boosted models. Discussing the differences between these algorithms is out of the scope of this study.}

% \myrevision{C5.0 is considered as the developed version of ID3 and C4.5, where it does not have any limitations in handling missing data, numerical attributes, and developing boosted models. Internal algorithm of CART is different than other methods, however, discussing technical details of these algorithms are beyond the scope of this study.}

% Here we use the C5.0 implementation \citep{Kuhn_2017_Manual}, which is an extension of the C4.5 algorithm developed by \citet{Quinlan_1993_Book, Quinlan_1996_JAIR}.

@@

The effectiveness of the tree, on the other hand, needs to be measured. To that end we use the factor $F_{\beta}$ proposed by \citet{Rijsbergen_1979_Book} as
% 
\begin{equation}
	\label{eq:f}
	F_{\beta} = \frac{ (1 + \beta^2) P R}{\beta^2 P + R}
	\, ,
\end{equation}
% 
where $P$ and $R$ are the precision and recall factors, respectively, and $\beta$ is a weighting factor between the two. $P$ measures the fraction of data samples that are correctly classified within a category, whereas $R$ measures the fraction of the data classified within that category regardless of whether it was correctly classified or not. Mathematically, they are defined as
% 
\begin{equation}
	P = \frac{ \mathrm{TP} }{ \mathrm{TP} + \mathrm{FP} } 
	\quad \mathrm{and} \quad
	R = \frac{ \mathrm{TP} }{ \mathrm{TP} + \mathrm{FN} }
	\, .
\end{equation}
% 
Here, TP, FP, and FN are the number of samples in the testing dataset considered to be true-positives, false-positives, and false-negatives, respectively. These values are typically ordered in what is known as a confusion matrix. For the simplest case of two categories, a confusion matrix takes the form
% 
\begin{equation}
\nonumber
\begin{array}{rr@{}rl}
	&					&	\multicolumn{2}{c}{\mathrm{Prediction}}		\\
	&					&	\mathrm{Positive}	&	\mathrm{Negative}	\\[0.75ex]
	\mathrm{Actual~Value}		
	&	\begin{array}{r}
			\mathrm{Positive} \\[1ex]
			\mathrm{Negative} 
		\end{array}
	&	\left[~
		\begin{array}{c}
			\mathrm{TP} \\[1ex]
			\mathrm{FP} 
		\end{array}\right.
	&
		\left.
		\begin{array}{cc}
			~\mathrm{FN} \\[1ex]
			~\mathrm{TN} \\[1ex]
		\end{array}~\right]
\end{array}
\end{equation}
% 
where TN is the number of true-negative samples. \change{These confusion matrices can be larger depending on the number of attributes. In our case, the confusion matrices are size 4$\times$4, in order to compare the actual number of samples with the predicted ones for the four given categories (poor, fair, good and excellent).}

### 2018-04-26 introduction.tex
close to finishing

Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation\change{, as well as a growing interest in the use of synthetic seismograms for engineering applications}. \change{Validation entails comparisons of simulations with observations, whereas verification involves comparisons of simulations with a exact or alternative solutions \citep[see][and references therein]{Bielak_2010_GJI, Taborda_2013_BSSA}}. Various methods or schemes have been proposed to evaluate, through direct quantitative comparisons or indirect statistical analysis, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}. \change{Some of these methods are better} suited for verification purposes because they compare the signals at the waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for validation. Among the latter, some are better for direct application purposes because they compare the behavior of systems (e.g., buildings or earth structures) at the response level \citep{Burks_BSSA_2014}, while the remaining methods focus on the validation of the overall characteristics of seismograms. In this category, the comparisons are done both in time and frequency. This type of validation is better suited for high-frequency or broadband simulations, where applications may be undefined and for which the expectation of matching seismograms at the waveform level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}). Within this group, methods such as those proposed by \citet{Anderson_2004_Proc} and \citet{Olsen_2010_SRL} are preferred because they offer simple schemes to quantify the goodness-of-fit (GOF) of synthetics with respect to data in a broad sense.

>>>>

Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation\change{, as well as a growing interest in the use of synthetic seismograms for engineering applications}. \change{Validation entails the comparison of simulations with observations, whereas verification involves the comparison of simulations with a exact or alternative solutions \citep[see][and references therein]{Bielak_2010_GJI, Taborda_2013_BSSA}}. Various methods or schemes have been proposed to evaluate, through direct \change{signal-to-signal} quantitative comparisons or \change{overall} statistical analysis, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}. \change{Some of these methods are better} suited for verification purposes because they compare the signals at the waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for validation. \change{\citet{Burks_BSSA_2014} and \citet{Rezaeian_2015_BSSA}, for instance, are used for validation in the context of engineering applications, i.e., at the level of the dynamic response of buildings. \citet{Anderson_2004_Proc} and \citet{Olsen_2010_SRL}, on the other hand, are used for validation in the context of the overall characteristics of seismograms when comparing synthetics with data in both the time and frequency domains. These methods are preferred for validating high-frequency or broadband simulations, where applications may be undefined and for which the expectation of matching seismograms at the waveform level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}), and because they offer simple schemes to quantify the goodness-of-fit (GOF) of simulation results.}

% Among the latter, some are better for direct application purposes because they compare the behavior of systems (e.g., buildings or earth structures) at the response level \citep{Burks_BSSA_2014}, while the remaining methods focus on the validation of the overall characteristics of seismograms. In this category, the comparisons are done both in time and frequency. This type of validation is better suited for high-frequency or broadband simulations, where applications may be undefined and for which the expectation of matching seismograms at the waveform level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}). Within this group, methods such as those proposed by \citet{Anderson_2004_Proc} and \citet{Olsen_2010_SRL} are preferred because they offer simple schemes to quantify the goodness-of-fit (GOF) of synthetics with respect to data in a broad sense.

### 2018-04-26 metrics.tex
close to finishing

There are various methods and algorithms available to evaluate the similarity between two or more seismograms, both through direct quantitative comparisons or indirect statistical analysis \citep[e.g.,][]{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}. In earthquake ground motion simulation, they are primarily used for verification with respect to benchmark or analytical solutions, and for validation with respect to data (i.e., ground motion records). Here, we focus on the list of metrics proposed by \citet{Anderson_2004_Proc}, with an additional metric for duration, as suggested by others \citep{Olsen_2010_SRL, Maufroy_2015_BSSA} and as implemented in \citet{Taborda_2013_BSSA}.

>>>>

There are various methods and algorithms available to evaluate the similarity between two or more seismograms, both through direct \chage{signal-to-signal} quantitative comparisons or \change{overall} statistical analysis \citep[e.g.,][]{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}. In earthquake ground motion simulation, they are primarily used for verification with respect to benchmark or analytical solutions, and for validation with respect to data (i.e., ground motion records). Here, we focus on the list of metrics proposed by \citet{Anderson_2004_Proc}, with an additional metric for duration, as suggested by others \citep{Olsen_2010_SRL, Maufroy_2015_BSSA} and as implemented in \citet{Taborda_2013_BSSA}.

### 2018-04-26 results.tex
close to finishing

In total, the clustering process results in 816, 1253, 879 and 76 data samples for the poor, fair, good and excellent classes, respectively. These groups are shown in Fig.~\ref{fig:count-classes}. As it can be seen in this figure, the number of samples in the excellent class is significantly less than those in the poor, fair, and good classes. Therefore, before moving on with the decision tree analysis, it is necessary to resample the subset of the excellent class. We used the oversampling approach described in the previous section. We replicated data from the excellent class 
\myrevision{through random selection of data from that class until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line. This oversampling is considered as a $\times$10 factor.} 
\myremove{randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.} 
% By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}. 
\change{According to \citet{Weiss_2003_JAIR}, an oversample rate of $\times$10 is considered to be acceptable in this type of data processing.}
(Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary. Instead, we used a \change{heavy} pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.

@@

For each one of these unique trees, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}), and extracted the total number of nodes in the trees and their depth. Figure \ref{fig:nodes-depth} shows the results of $F_1$ for all the trees and its distribution in terms of the number of attributes used in the trees as a function of the number of nodes and the depth of each tree. Recall that we are interested in finding a sequence of decisions (represented by disjunctive decision nodes in a tree) that can lead to good GOF predictions (i.e, high values of $F_1$) using a reduced number of attributes (GOF metrics). In general, all the trees obtained with the C5.0 algorithm are good in terms of the effectiveness factor ($F_1$ values close to 1). Then our choice comes down to using a reduced number of metrics. Having several trees with 2, 3, and 4 metrics (as opposed to 11), the following factor in the decision is choosing trees with algorithms using a small number of steps to reach the prediction. This is given by a combination between the depth of the tree and the number of nodes in the tree (i.e., trees with low complexity).

>>>>

In total, the clustering process results in 816, 1253, 879 and 76 data samples for the poor, fair, good and excellent classes, respectively. These groups are shown in Fig.~\ref{fig:count-classes}. As it can be seen in this figure, the number of samples in the excellent class is significantly less than those in the poor, fair, and good classes. Therefore, before moving on with the decision tree analysis, it is necessary to resample the subset of the excellent class\change{, for which we use the oversampling approach described in the previous section. Oversampling is nothing else but a replication of data samples. This process is done randomly, i.e., through a random selection of data samples from the original set to be duplicated until one increases the number of total samples in the oversampled set to a desired target number. In the case of the excellent class, we applied oversampling until reaching a total of 760 samples, as indicated with the dashed line in Fig.~\ref{fig:count-classes}. Since the original size was 76 samples, that means we applied an oversampling ratio of 10$\times$. According to \citet{Weiss_2003_JAIR}, an oversample rate of 10$\times$ is considered to be acceptable in this type of data processing.} (Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary. Instead, we used a \change{heavy} pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.

% We used the oversampling approach described in the previous section. We replicated data from the excellent class 
% \myrevision{through random selection of data from that class until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line. This oversampling is considered as a $\times$10 factor.} 
% \myremove{randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.} 
% By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}.

@@

For each one of these unique trees, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}), and extracted the total number of nodes in the trees and their depth. Figure \ref{fig:nodes-depth} shows the results of $F_1$ for all the trees and its distribution in terms of the number of \change{decision} attributes \change{or GOF metrics} used in the trees as a function of the number of nodes and the depth of each tree. Recall that we are interested in finding a sequence of decisions (represented by disjunctive decision nodes in a tree) that can lead to good GOF predictions (i.e, high values of $F_1$) using a reduced number of attributes (GOF metrics). In general, all the trees obtained with the C5.0 algorithm are good in terms of the effectiveness factor ($F_1$ values close to 1). Then our choice comes down to using a reduced number of metrics. Having several trees with 2, 3, and 4 metrics (as opposed to 11), the following factor in the decision is choosing trees with algorithms using a small number of steps to reach the prediction. This is given by a combination between the depth of the tree and the number of nodes in the tree (i.e., trees with low complexity).
